---
title: "Exercise week 5"
author: "Aloïs THOMAS"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(AnnotationHub)
  library(ensembldb)
  library(GenomicRanges)
  library(epiwraps)
  library(rtracklayer)
  library(ggplot2)
})
ah <- AnnotationHub()
```
## Part 1
# Loading peaks
```{r}
#H3K4me3
download.file("https://www.encodeproject.org/files/ENCFF331RHM/@@download/ENCFF331RHM.bed.gz", "H3K4me3.bed.gz", mode = "wb")
H3K4me3_peaks<- rtracklayer::import("H3K4me3.bed.gz", format="narrowPeak")
#H3K27me3
download.file("https://www.encodeproject.org/files/ENCFF105NKG/@@download/ENCFF105NKG.bed.gz", "H3K27me3.bed.gz", mode = "wb")
H3K27me3_peaks<- rtracklayer::import("H3K27me3.bed.gz", format="narrowPeak")
```
#Length of Bivalent domains
```{r}
bivalent_domains <- H3K4me3_peaks[overlapsAny(H3K4me3_peaks, H3K27me3_peaks)]

bival_length <- length(bivalent_domains)
bival_length

```


## Part 2
# Peaks loading of smooth muscle cells

```{r}
download.file("https://www.encodeproject.org/files/ENCFF817CHW/@@download/ENCFF817CHW.bed.gz", dest="H3K27me3.bed.gz")
H3K27me3_peaks_sm<- rtracklayer::import("H3K27me3.bed.gz", format="narrowPeak")
download.file("https://www.encodeproject.org/files/ENCFF714BRC/@@download/ENCFF714BRC.bed.gz", dest="H3K4me3.bed.gz")
H3K4me3_peaks_sm<- rtracklayer::import("H3K4me3.bed.gz", format="narrowPeak")

```
#Overlaps calculation between sm H3K4me3, H3K27me3 and mESCs
```{r}
overlaps <- list(bivalent_domains, H3K27me3_peaks_sm, H3K4me3_peaks_sm)
names(overlaps) <- c("overlap_ESC", "overlap_H3K27me3_sm", "overlap_H3K4me3_sm")
regionUpset(overlaps[c(1,2,3)])
```

From the plot, we can see the following:

52427 H3K27me3 sm domains, 26397 H3K4me3 domains and 2934 mESC domains which do not overlap between each other.

22 domains are overlapping each other (mESCs and Sms)

52 domains overlapping for mESC and H3K4me3 sm

70 domains overlapping for mESC and H3K27me3 sm

4237 domains overlapping for H3K27me3 sm and H3K4me3 sm 





