---
title: "assignment"
author: "Aloïs THOMAS"
output: html_document
---


```{r}
suppressPackageStartupMessages({
  library(GenomicRanges)
  library(epiwraps)
  library(ggplot2)
  library(rGREAT)
})
```



```{r}
options(timeout = 1000)
download.file("https://ethz-ins.org/content/w10.assignment.zip", "w10.assignment.zip")
unzip("w10.assignment.zip")
```

## Prepare tracks/regions

```{r}
tracks <- list.files(pattern="bw$")
peaks <- list.files(pattern="bed$")
# we first import the peaks
peaks <- lapply(peaks, rtracklayer::import.bed)
# we'll focus on the high-quality peaks
peaks <- lapply(peaks, FUN=function(x) x[x$score>800])
# we get the union of non-redundant regions
regions <- reduce(unlist(GRangesList(peaks)))
```

## Plot the signals

```{r}
ml <- signal2Matrix(tracks, regions, extend=2000)
plotEnrichedHeatmaps(ml)
```

## Clustering

```{r}
k=4
set.seed(123)
cl <- clusterSignalMatrices(ml, k=k)
mycolors <- scales::hue_pal()(k)
names(mycolors) = 1:k
plotEnrichedHeatmaps(ml, row_split=cl, mean_color=mycolors)
d <- meltSignals(ml, splitBy=cl)
#plot averages
ggplot(d, aes(position, mean, colour=sample)) + geom_line() + facet_wrap(~split)
```

Cluster interpretation:

- Cluster 1: bound mainly by CREB1, also a but of CREB3L1

- Cluster 2: bound by CREB1, CREB3, CREB3L1.

- Cluster 3: bound by CREB3

- Cluster 4: bound by CREB1 and CREB3L1

Thus, CREB1 and CREB3L1 are bound often together to the same region, while CREB3 moslty bind solely alone on its region (cluster 3).


## Enrichment analysis (focus on cluster 3)

```{r}
# we first split the regions by cluster:
split_regions <- split(regions, cl)
# we send the data to the GREAT server to run an enrichment analysis
#Focus on cluster 3
job <- submitGreatJob(gr=split_regions[["3"]], bg=regions, species="hg38")
res <- getEnrichmentTables(job)
```

Plot top molecular functions:

```{r, fig.width=9, fig.height=6}
bp <- res$`GO Molecular Function`
head(bp)
ggplot(head(bp,15), aes(Hyper_Fold_Enrichment, reorder(name, Hyper_Adjp_BH), size=Hyper_Foreground_Region_Hits, color=-log10(Hyper_Adjp_BH))) + geom_point() + scale_color_viridis_c()
```


The enrichment analysis for cluster 3 reveals that these regions are related to genes involved in receptor binding and activity. These receptors are mainly related to cytokines and involved in some cell death pathways.